---
title: "progress-report-kushpamnani.Rmd"
author: "Kush Pamnani"
date: "2025-04-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries
```{r libraries}
library(rvest)
library(dplyr)
library(readr)
library(stringr)
```

## Scrape the Hall of Fame 2025 Data
```{r scrape-data}
url <- "https://www.baseball-reference.com/awards/hof_2025.shtml"
html <- read_html(url)
tables <- html_table(html)

# Assign the first table to a data frame
data <- tables[[1]]
```

## Clean the Data
```{r clean-data}
# The first row has actual column names
actual_col_names <- data[1, ]
colnames(data) <- actual_col_names

# Force unique column names to avoid duplicates
colnames(data) <- make.names(colnames(data), unique = TRUE)

# Remove the first row which was header information
data <- data[-1, ]

# Convert Votes and %vote columns to numeric
data$Votes <- as.numeric(data$Votes)
data$X.vote <- parse_number(data$X.vote)
```

## Prepare Final Dataset to Match Lahman's HallOfFame Structure
```{r prepare-final-data}
# Assume total ballots is 394
total_ballots <- 394
needed_votes <- ceiling(0.75 * total_ballots)

# Create the final dataset
final_data <- data %>%
  mutate(
    playerID = NA, # Manual matching needed normally
    yearID = 2025,
    votedBy = "BBWAA",
    ballots = total_ballots,
    needed = needed_votes,
    votes = Votes,
    inducted = ifelse(X.vote >= 75, "Y", "N"),
    category = "Player",
    needed_note = NA
  ) %>%
  select(playerID, yearID, votedBy, ballots, needed, votes, inducted, category, needed_note)

# View the cleaned final dataset
head(final_data)
```

## Save Individual Work (Optional)
```{r save-final}
# Save your individual work if needed (not required)
# write_csv(final_data, "hall_of_fame_2025_individual.csv")
```

## Notes
- Successfully scraped the 2025 Hall of Fame voting data from Baseball Reference.
- Cleaned column names and removed duplicates.
- Converted necessary columns to numeric.
- Prepared the final dataset to match the structure expected in the Lahman HallOfFame table.
